Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}

### 2021-01-25 main.tex
tweak

Fix pairwise disjoint countable sets $\VAR=\set{x,y,z,\ldots}$ of variables,
$\FUN=\set{f,g,\ldots}$ of function symbols, and 
$\PRED=\set{p,q,\ldots}$ of predicate symbols. Each function symbol and
each predicate symbol has a finite arity. The grammars below generate
the set $\TERM$ of \bfit{terms}, denoted by $s,t,u,\ldots$, the set
$\ATOM$ of \bfit{atoms}, denoted by $a,b,c,\ldots$, and the set
$\FORM$ of \bfit{formulas}, denoted by $A,B,C,\ldots$:
\begin{equation*}
  \begin{array}{r@{~}l}
    t &\coloneqq~ x \mid f(t_1,\dots,t_n)
    \\[.5ex]
    a &\coloneqq~ \ttt\mid\fff\mid p(t_1,\dots,t_n) \mid \dual p(t_1,\dots,t_n)
    \\[.5ex]
    A &\coloneqq~ a \mid A\vlan A \mid A\vlor A \mid \exists x.A \mid \forall x.A
\end{array}
\end{equation*}
where the arity of $f$ and $p$ is $n$.
For better readability of often omit parentheses and write simply $ft_1 \dots t_n$ or $pt_1 \dots t_n$.
We consider the truth constants
$\ttt$ (\emph{true}) and $\fff$
(\emph{false}) as additional atoms, and we consider all formulas in negation
normal form, where \bfit{negation} $(\dual\cdot)$ is defined on
atoms and formulas via De Morgan's laws:
\begin{equation*}
  \begin{array}{ccc}
    \dual{\dual a} = a
    &
  \dual\ttt =\fff &
  \dual{p(t_1,\ldots,t_n)} = \dual{p}(t_1,\ldots,t_n) \\[.5ex]
  &
  \dual\fff = \ttt &
  \dual{\dual p(t_1,\ldots,t_n)} = p(t_1,\ldots,t_n) \\[.5ex]
  &
  \dual{\exists x.A} = \forall x.\dual{A}
  &
  \dual{A \vlan B} = \dual{A} \ \vlor \ \dual{B} 
  \\[.5ex]
  &
  \dual{\forall x.A} = \exists x.\dual{A}
  &
%  \\[.5ex]
  \dual{A \vlor B} = \dual{A} \ \vlan \ \dual{B}
  \end{array}
\end{equation*}
Then we write $A\implies B$ as abbreviation for $\dual A\vlor B$.

>>>>

Fix pairwise disjoint countably infinite sets $\VAR=\set{x,y,z,\ldots}$ of variables,
$\FUN=\set{f,g,\ldots}$ of function symbols, and 
$\PRED=\set{p,q,\ldots}$ of predicate symbols. Each function symbol and
each predicate symbol has a finite arity. The grammars below generate
the set $\TERM$ of \bfit{terms}, denoted by $s,t,u,\ldots$, the set
$\ATOM$ of \bfit{atoms}, denoted by $a,b,c,\ldots$, and the set
$\FORM$ of \bfit{formulas}, denoted by $A,B,C,\ldots$:
\begin{equation*}
  \begin{array}{r@{~}l}
    t &\coloneqq~ x \mid f(t_1,\dots,t_n)
    \\[.5ex]
    a &\coloneqq~ \ttt\mid\fff\mid p(t_1,\dots,t_n) \mid \dual p(t_1,\dots,t_n)
    \\[.5ex]
    A &\coloneqq~ a \mid A\vlan A \mid A\vlor A \mid \exists x.A \mid \forall x.A
\end{array}
\end{equation*}
where the arity of $f$ and $p$ is $n$.
For better readability of often omit parentheses and write simply $ft_1 \dots t_n$ or $pt_1 \dots t_n$.
We consider the truth constants
$\ttt$ (\emph{true}) and $\fff$
(\emph{false}) as additional atoms, and we consider all formulas in negation
normal form, where \bfit{negation} $(\dual\cdot)$ is defined on
atoms and formulas via De Morgan's laws:
\begin{equation*}
  \begin{array}{ccc}
    \dual{\dual a} = a
    &
  \dual\ttt =\fff &
  \dual{p(t_1,\ldots,t_n)} = \dual{p}(t_1,\ldots,t_n) \\[.5ex]
  &
  \dual\fff = \ttt &
  \dual{\dual p(t_1,\ldots,t_n)} = p(t_1,\ldots,t_n) \\[.5ex]
  &
  \dual{\exists x.A} = \forall x.\dual{A}
  &
  \dual{A \vlan B} = \dual{A} \ \vlor \ \dual{B} 
  \\[.5ex]
  &
  \dual{\forall x.A} = \exists x.\dual{A}
  &
%  \\[.5ex]
  \dual{A \vlor B} = \dual{A} \ \vlan \ \dual{B}
  \end{array}
\end{equation*}
Then we write $A\implies B$ as abbreviation for $\dual A\vlor B$.
